options(scipen = 999) # turns of scientific notations like 1e+40
library(ggplot2)
library(ggthemes)
library(dplyr)
library(lubridate)
library(pipeR)
library(data.table)
library(pacman)
library(tidyr)
pacman::p_load(tm,SnowballC,foreign,plyr, twitteR,slam,foreign,wordcloud,LiblineaR,e1071,caret,ranger,rpart)
detach("package:plyr", unload=TRUE) 

##Descriptive stats
des = read.csv ("/Users...des_no_outliers.csv", stringsAsFactors = F)
des_outlier = read.csv ("/Users...des.csv", stringsAsFactors = F)
des = des[,-which(names(des) == "X")] #removing "X"
des_outlier = des_outlier[,-which(names(des_outlier) == "X")] #removing "X"

attach(des)
```


Summary table

```{r}
dim(des)
table(des$country)
table(des$success)
summary(des$funding_total_usd)
summary(des$funding_rounds)
summary(des$last_first_funding)
table(des$success)

aggregate(data.frame(count = year), list(value = year), length)
aggregate(data.frame(count = region), list(value = region), length)
aggregate(data.frame(count = country), list(value = country), length)
aggregate(data.frame(count = country), list(value = country), length)

```


Graphs of USD funding total
```{r}
summary(des$funding_total_usd)
summary(des_outlier$funding_total_usd)

ggplot(aes(x = funding_total_usd/1000000), data = des) +
  geom_histogram(binwidth = .08) + 
  scale_x_log10() +
  xlab('Million USD') +
  ggtitle('Histogram of the Raised Amount') +
  theme(plot.title = element_text(face = "italic")) 

ggplot(aes(x = funding_total_usd/1000000, fill=success), data = des) +
  geom_histogram(binwidth = .09) + 
  scale_x_log10() +
  xlab('Million USD') +
  ggtitle('Histogram of the Raised Amount') +
  theme(plot.title = element_text(face = "italic")) 

ggplot(data=des, aes(x=funding_rounds, y=funding_total_usd/1000000, colour = success)) + geom_point(alpha = 0.3,  position = position_jitter()) + xlab('Funding round') + ylab('Total Funding (millions)')
dev.copy(pdf,'rounds+total funding.pdf')
dev.off()



des %>%
  filter(success == "0") %>%
  ggplot(aes(x = funding_rounds, y = funding_total_usd, colour = success)) +
  geom_point(alpha = 0.3,  position = position_jitter()) + stat_smooth(method = "lm")

des %>%
  filter(success == "1") %>%
  ggplot(aes(x = funding_rounds, y = funding_total_usd, colour = success)) +
  geom_point(alpha = 0.3,  position = position_jitter()) + stat_smooth(method = "lm")
```

## Top 10 Funding Total raised by Company

```{r}
des <- des [order(funding_total_usd, decreasing = TRUE), ]
funding <- head(des, 10)

ggplot(aes(x = reorder(name, funding_total_usd), y = funding_total_usd/1000000), data = funding) +
  geom_bar(stat = 'identity') +
  xlab('Startups') +
  ylab('Million USD') +
  ggtitle('Top 20 Total Funding Raised by a Startup') + coord_flip() + theme(plot.title = element_text(face = "italic"))



theme_set(theme_bw())
# Plot
ggplot(funding, aes(x= reorder(name,funding_total_usd), y=funding_total_usd/1000000)) + 
  geom_point(size=3) + 
  geom_segment(aes(x=name, 
                   xend=name, 
                   y=0, 
                   yend=funding_total_usd/1000000)) + 
  labs(title="Top 10 Startups by Funding") +
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
dev.copy(pdf,'top10_startups.pdf')
dev.off()

```

## Top 20 Funding Total raised by region

```{r}
group.region.1 <- group_by(des, region, success)
#group.region <- summarize(group.region.1, funding_total_usd = n())
funding_by_region <- summarize(group.region.1, funding_sum = sum(funding_total_usd))
funding_by_region <- funding_by_region[order(funding_by_region$funding_sum, decreasing = TRUE), ]
funding_by_region_top20 <- head(funding_by_region, 10)

ggplot(aes(x = reorder(region, funding_sum), y = funding_sum/1000000000), data = funding_by_region_top20) +
  geom_bar(stat = 'identity') +
  xlab('Regions') +
  ylab('Billion USD') +
  ggtitle('Top 10 Regions Funding Raised by Startups') + coord_flip() + theme(plot.title = element_text(face = "italic"))

dev.copy(pdf,'funding_by region.pdf')
dev.off()

```

## Total Funding raised by year

```{r}
#fix the data to get year
des$founded_at <- strptime(des$founded_at, "%m/%d/%y") %>>% as.POSIXct()
des$year = format(as.Date(des$founded_at, format="%d/%m/%Y"),"%Y")
des$year[is.na(des$year)] = "other_year"
table(des$year)
des = des[ which(des$year>="2000" & des$year<"2016"), ] #keep only from year 2000 to 2017

#group the years by funding
group.year.1 <- group_by(des, year)
funding_by_year <- summarize(group.year.1, funding_sum = sum(funding_total_usd))

#group the years worldwide
ggplot(funding_by_year, aes(x = year, y=funding_sum/1000000000)) + geom_bar(stat="identity") +  
  ylab("Billions of USD in Investments (World)") + xlab("Year") + theme(plot.title = element_text(face = "italic"))

dev.copy(pdf,'years_fund_world.pdf')
dev.off()


#group the years in SF bay area
group.year.2 <- group_by(des, year, region)
funding_by_year.2 <- summarize(group.year.2, funding_sum = sum(funding_total_usd))
funding_by_year.2 <- subset(funding_by_year.2, region == "SFBayArea") #keep only "SF" 

ggplot(funding_by_year.2, aes(x = year, y=funding_sum/1000000000)) + geom_bar(stat="identity") +  
  ylab("Billions of USD in Investments (SF Bay Area)") + xlab("Year") + theme(plot.title = element_text(face = "italic"))

dev.copy(pdf,'years_fund_SF.pdf')
dev.off()

```
## Top 20 Funding Total raised by region and industry

```{r}
group.region <- group_by(des, category_list)
funding_by_region <- summarize(group.region, funding_sum = sum(funding_total_usd))
funding_by_region <- funding_by_region[order(funding_by_region$funding_sum, decreasing = TRUE), ]
funding_by_region_top20 <- head(funding_by_region, 10)
funding_by_region_top20 [funding_by_region_top20 == "Biotechnology|Health Care"] = "Biotechnology + Health Care"
funding_by_region_top20 [funding_by_region_top20 == "Internet|Mobile"] = "Internet + Mobile"
funding_by_region_top20 [funding_by_region_top20 == "All Students|Colleges|Communities|Identity|Social Media"] = "Communities + Social Media"


ggplot(aes(x = reorder(category_list, funding_sum), y = funding_sum/1000000000), data = funding_by_region_top20) +
  geom_bar(stat = 'identity') +
  xlab('Industry in the World') +
  ylab('Billion USD') +
  ggtitle('Top 10 Industries by Funds Raised') + coord_flip() + theme(plot.title = element_text(face = "italic"))
dev.copy(pdf,'world+industry+funds.pdf')
dev.off()
```

## Success rate in SF area by industry from top 20 funding 

```{r}
sf_bay <- subset(des, region == "SFBayArea") #keep only "SF" 
group.region <- group_by(sf_bay, category_list)
funding_by_region <- summarize(group.region, funding_sum = sum(funding_total_usd))
funding_by_region <- funding_by_region[order(funding_by_region$funding_sum, decreasing = TRUE), ]
funding_by_region_top20 <- head(funding_by_region, 10)

funding_by_region_top20 [funding_by_region_top20 == "All Students|Colleges|Communities|Identity|Social Media"] = "Communities + Social Media"
funding_by_region_top20 [funding_by_region_top20 == "Biotechnology|Health Care"] = "Biotechnology + Health Care"
funding_by_region_top20 [funding_by_region_top20 == "Information Services|Messaging|MicroBlogging|Service Providers|SMS|Software"] = "Information Services + Messaging + Software"
funding_by_region_top20 [funding_by_region_top20 == "Clean Technology|Construction"] = "Clean Technology + Construction"

ggplot(aes(x = reorder(category_list, funding_sum), y = funding_sum/1000000000), data = funding_by_region_top20) +
  geom_bar(stat = 'identity') +
  xlab('Industry in SF Bay Area') +
  ylab('Billion USD') +
  ggtitle('Top 10 Industries by Funds Raised') + coord_flip() + theme(plot.title = element_text(face = "italic"))
dev.copy(pdf,'SF+industry+funds.pdf')
dev.off()

```


